---
title: "Association Rules "
author: " "
date: "July 21, 2017"
output: html_document
---
# Problem 3 - Assignment
### Association Rules for "Flight Delay" dataset: Generate the rules and identify the patterns. 

1. Read the data into R

```{r, echo=T}
rm(list=ls(all=T))

```


2. Look at the summary of all the variables and convert the following variables as factors 
        + Weather 
        + DAY_WEEK 
        + Flight Status
```{r}
setwd("C:/insofe\\lab_assignments\\Association_Rules_Assignment")
getwd()
flight_data=read.csv("FlightDelays.csv")
str(flight_data)
flight_data_int=c("Weather","DAY_WEEK","Flight.Status")
flight_data[,flight_data_int]=data.frame(apply(flight_data[,flight_data_int],2,function(x) as.factor(as.character(x))))
str(flight_data)
```

3. Bin the numeric variable 'CRS_DEP_TIME' into 4 bins as follows: If time is between 6 AM to 12 noon code as 1 and If the time is between 2 , etc.
```{r}
library(infotheo)
library(arules)
flight_data[,1]=discretize(flight_data[,1], method="interval", categories = 4, labels =c(1:4) )
str(flight_data)
```

4. Convert the data frame in a transactions object. Look at the first 6 transactions to understand how 'arules' library requires data to be transformed. Use the below R code to view the transactions data 
```{r, echo=T}

flight_trans <- as(flight_data, "transactions")
inspect(flight_trans[1:6])
flight_trans
image(flight_trans)
itemFrequency(flight_trans)
itemFrequencyPlot(flight_trans)
```

5. Apply 'arules' algorithm and play with various support, lift and confidence values.
```{r}
rules <- apriori(flight_trans,parameter = list(sup = 0.05, conf = 0.6,target="rules"))
summary(rules)
#rules <- apriori(flight_trans,parameter = list(sup = 0.1, conf = 0.7,target="rules"))
#summary(rules)
#rules3 <- apriori(flight_trans,parameter = list(sup = 0.2, conf = 0.8,target="rules"))
#summary(rules3)
```

6. Inspect all the rules.
```{r, echo=T}
inspect(rules)
library(arulesViz)
plot(rules)
```
```
7. Filter the rules with specific LHS and RHS conditions 
E.g.; Filter the rules with Flighstatus=0
```{r}

rules.itemfilter1=as(subset(rules, subset = rhs %in% "Flight.Status=0"),"data.frame")
rules.itemfilter1 

rules.itemfilter3=as(subset(rules, subset = rhs %in% "Flight.Status=0"),"data.frame")
rules.itemfilter3 

```
```
8. Filter redundant rules if any
```{r, echo=T}
rules.sorted =sort(rules, by="lift")
subset.matrix =is.subset(rules.sorted, rules.sorted,sparse = FALSE)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant = colSums(subset.matrix, na.rm=T) >= 1
which(redundant)
rules.pruned =rules.sorted[!redundant]
inspect(rules.pruned)
```


9. Pot and visualize the rules
```{r}
library(arulesViz)
plot(rules.pruned)
```